---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# Spurious correlations

<!-- badges: start -->
<!-- badges: end -->

The goal of spuriouscorrelations is to keep alive the amazing examples from
[Tyler Vigen](https://web.archive.org/web/20230607181247/https://tylervigen.com/spurious-correlations).
Unfortunately, as of 2023-10-09, the website is down as my students noticed.
Therefore, I decided to use the snapshot from the Internet Wayback Machine to
save the datasets from 2023-06-07.

## Installation

You can install the development version of spuriouscorrelations like so:

``` r
remotes::install_github("pachadotdev/spuriouscorrelations")
```

## Example

This is a basic example which shows you how to plot a spurious correlation:

```{r example}
library(dplyr)
library(ggplot2)
library(spuriouscorrelations)

unique(spurious_correlations$var2_short)

nicholas_cage <- spurious_correlations %>%
  filter(var2_short == "Nicholas Cage")

ggplot(nicholas_cage) +
  geom_point(aes(x = var1_value, y = var2_value, color = year), size = 4) +
  theme_minimal() +
  labs(
    x = "Drownings by falling into a pool",
    y = "Nicholas Cage movies",
    title = "Drownings by Falling into a Pool vs Nicholas Cage Movies"
  )
```

The correlation is:
```{r example2}
cor(nicholas_cage$var1_value, nicholas_cage$var2_value)
```

Now let's make it double-axis:
  
```{r example3}
library(tidyr)

nicholas_cage_long <- nicholas_cage %>%
  select(year, var1_value, var2_value) %>%
  pivot_longer(
    cols = c(var1_value, var2_value),
    names_to = "variable",
    values_to = "value"
  ) %>%
  # standardize the values
  group_by(variable) %>%
  mutate(
    value = (value - mean(value)) / sd(value),
    variable = case_when(
      variable == "var1_value" ~ "Drownings by falling into a pool",
      variable == "var2_value" ~ "Nicholas Cage movies"
    )
  )

# make a double y axis plot with year on the x axis
ggplot(nicholas_cage_long, aes(
  x = year, y = value, color = variable,
  group = variable
)) +
  geom_line() +
  geom_point() +
  theme_minimal() +
  theme(legend.position = "top") +
  labs(
    x = "Year",
    y = "Standardized values",
    title = "Drownings by Falling into a Pool vs Nicholas Cage Movies"
  )
```
